Cartogram Mapping and its Application to Cancer Data Visualisation

### 2019-09-15 review_article.Rmd
Table of caption information for USA plots

<!-- what is a choropleth map -->
A choropleth map is used to display the spatial relationship of measurements by shading areas of a map. The geography is faithfully rendered, and the colour rendering is designed to reveal spatial patterns among data values. A choropleth is constructed by drawing the geographic or political boundaries, and filling the shapes with colours to represent values of a measured variable [@EI]. 
Figure \ref{fig:choro} shows a choropleth of age-adjusted rate (per 100,000 people) of new cases of lung and bronchus in the USA, averaged over 2012 through 2016. The data was extracted from the [official federal statistics](https://gis.cdc.gov/cancer/USCS/DataViz.html) on cancer incidence and deaths, produced by the Centers for Disease Control and Prevention (CDC) and the National Cancer Institute (NCI).

@@

b <- st_bbox(cancer)
ggplot(cancer) + 
  geom_sf(aes(fill = AgeAdjustedRate)) +
  scale_fill_distiller("Age-adjusted rate \n (per 100,000)", type = "seq", palette = "YlOrRd",  direction = 1) + 
  #coord_sf(crs = CRS("+init=epsg:3857"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
  theme_void() +
  theme(legend.position ="bottom", legend.key.width = unit(2, "cm"))
```

@@

@@

<!-- What is a contiguous cartogram -->

<!-- Intentionally preserve neighbours -->

>>>>

<!-- what is a choropleth map -->
A choropleth map is used to display the spatial relationship of measurements by shading areas of a map. The geography is faithfully rendered, and the colour rendering is designed to reveal spatial patterns among data values. A choropleth is constructed by drawing the geographic or political boundaries, and filling the shapes with colours to represent values of a measured variable [@EI]. 
Figure \ref{fig:choro} shows a choropleth of age-adjusted rate (per 100,000 people) of new cases of lung and bronchus in the USA, averaged over 2012 through 2016. The data was extracted from the official federal statistics by @usa_lung on cancer incidence and deaths, produced by the Centers for Disease Control and Prevention (CDC) and the National Cancer Institute (NCI).

@@

b <- st_bbox(cancer)
ggplot(cancer) + 
  geom_sf(aes(fill = AgeAdjustedRate), colour = NA) +
  scale_fill_distiller("Age-adjusted rate \n (per 100,000)", type = "seq", palette = "RdPu",  direction = 1) + 
  #coord_sf(crs = CRS("+init=epsg:3857"), xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"])) +
  theme_void() +
  theme(legend.position ="bottom", legend.key.width = unit(2, "cm"))
```

@@

```{r gghexmap, fig.cap = "A hexagon tesselation map of the Unites States of America. Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016. Each state has been represented by a square, each square is the same size. The neighbouring states are easily contrasted, however the north east regions has been displaced from their geographic location. The sparsity of the population in the Mid West is highlight by the light yellow colour, the age adjusted rate in Kentucky is the darkest and its neighbours are similar.", message=FALSE, warning=FALSE, out.width='80%'}
knitr::include_graphics("figures/gghexmap.png")
```

Table: \label{tab:usa} Maps used to present statistics for the Uited States of America

|Map display |Details| 
|:--------------------|:------------------------------------------------|  
|a. contiguous cartogram|A contiguous cartogram map of the United States of America. Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016. Each stated shape has been distorted according to the population of the state in 2015. The state of California has become much larger due to it's large population density. This draws attention to the densely populated North East region, and detracts from the less populated Mid West.|
|b. Non - Contiguous|A Non - contiguous cartogram map of the United States of America. Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016. Each state shape has been maintained, but the size has altered according to the population of the state in 2015. The state of California has remained closer to it's original size than it's surrounding states. The North East states have remained closer to their geographical size, in the case of Massachusetts and Connecticut. This draws attention to the densely populated North East region, and the sparse Mid West.|
|Dorling|A dorling cartogram of the Unites States of America. Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016. Each state has been represented by a circle, but the size was determined by the population of the state in 2015. The North East states remain closer to their neighbours, and may be displaced from their geographic location. The sparsity of the population in the Mid West is highlight by the distance between the circles, located at the geographic centroids.|
|Hexagon Tesselation|A hexagon tesselation map of the Unites States of America. Each state has been coloured according to the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016. Each state has been represented by a square, each square is the same size. The neighbouring states are easily contrasted, however the north east regions has been displaced from their geographic location. The sparsity of the population in the Mid West is highlight by the light yellow colour, the age adjusted rate in Kentucky is the darkest and its neighbours are similar.|

@@

<!-- What is a contiguous cartogram -->
<!-- Intentionally preserve neighbours -->
